---
title: Using gDRstyle
author:
- name: Allison Vuong 
  email: gdr-support-d@gene.com
package: gDRstyle
date: 
output:
  BiocStyle::html_document:
    toc_float: yes
vignette: >
  %\VignetteIndexEntry{gDRstyle}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE}
library(BiocStyle)
library(gDRstyle)
```

# Overview

The `gDRstyle` package is intended to be used during development of packages 
within the gDR platform. It has 3 primary uses: (1)to set a style guide with 
functions that check that the style is upheld, (2) during CI to ensure code 
passes `R CMD check` to maintain the state of the code in high quality, and (3) 
for package dependency installation during gDR platform image building.

# Use Cases

## Style guide
See the written [Style guide](https://github.com/gdrplatform/gDRstyle/blob/gh-pages/style_guide.md). 
The function `lintPkgDirs` can be used to ensure the package is appropriately 
linted. 
## CI/CD
The `checkPackage` function will check that the package abides by gDRstyle 
stylistic requirements, passes `rcmdcheck`, and ensures that the 
`dependencies.yml` file used to build gDR platform's docker image is kept 
up-to-date with the dependencies listed in the package's `DESCRIPTION` file. 
This is called in gDR platform packages' CI/CD. 

## Package installation
The function `installAllDeps` assists in installing package dependencies.  

# SessionInfo {-}

```{r sessionInfo}
sessionInfo()
```
