# gDR style

## General

1. Indentation
  * 2 spaces for indentation
2. Use of spaces
  * Always utilize `if (` over `if(`
  * Always include spaces around logical comparison `1 == 1` over `1==1`
  * Always include spaces around logical entity `if (1 == 1) {}` over `if (1=1){}`
  * Always include spaces around "full" subsets `[ , subset]` over `[, subset]`
3. Naming conventions
  * variables
    * Utilize lower_snake_case for variables
    * Assignment should use `<-` over `=`
  * function arguments
    * Utilize space between function arguments `function(a = "A")` over `function(a="A")`
    * Utilize lower_snake_case for function arguments `function(big_matrix)` over `function(bigMatrix)`
  * function names
    * All action functions should start with verbs `compute_metrics_SE` over `metrics_SE`
    * All internal functions start with a `.`
    * Utilize lowerCamelCase for function names
    * Function names should be short and informative
    * Getters should start with `get*`, setters should start with `set*`, boolean checkers should start with `is*`
4. Functions, functions, functions
  * Functions definition
    * Functions definition should start on the same line as function name
    * Every parameter should be in separate line
    * Prepare separate functions for complicated defaults (of function parameters)
```{r}
 # Good
fun <- function(param1,
                param2,
                param_with_dir_for_st_important = get_st_important_dir()) {
  # Code is indented by two spaces.
  ...
}
 # Bad
fun <- function(param1, param2, param_with_dir_for_st_important = file.path(system.file(paste(param1, "SE", "rds", sep = "."), package = "important_package"))) {
  ...
}

```

  * Function assignment
    * Always utilize `<-` over `=` to differentiate function arguments assignments from function assignments `myFunction <- function()` over `myFunction = function()`
  * Function export
    * All internal functions are not exported
    * All exported functions have `assert` tests for their parameters
  * `vapply` over `sapply` (or `lapply` + `unlist()` if predefining FUN.VALUE is difficult)
  * Do not use `apply` on data.frame(s) (`mapply` is good for row-wise operations) 
6. Use of curly braces
  * `if` and `else` statements should be surrounded by curly braces on the same line

```{r}
if (TRUE) {
  NULL
} else {
  NULL
}
```

  * Assign simple if-else statements to a variable.
```{r}
what_is_going_on <- if (is_check()) {
  flog <- "it's getting hot..."
} else if (is_mate()) {
  flog <- "Oh noooo..."
} else {
  flog <- "there is a hope..."
}
```

8. Files
  * Where possible, break code into smaller files
  * Separate each function by 2 lines
  * Tests should be named identically to the file in `R/` (`R/assays` => `tests/testthat/test-assays.R`)
9. Miscellaneous
  * Exponentiation: always utilize `^` over `**` for exponentiation like `2 ^ 3` over `2**3`.
